### 2023-03-16 paper.tex
Initial final content

@@

\begin{abstract}
  The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
  both left and right-hand margins. Use 10~point type, with a vertical
  spacing of 11~points.  The {\bf Abstract} heading must be centered,
  bold, and in point size 12. Two line spaces precede the
  Abstract. The Abstract must be limited to one paragraph.
\end{abstract}

@@

\subsection{Attention masking}

TODO: explain how attention masking works

TODO: explain sparse transformer, used in gpt-3 (maybe 2 also)

The original transformer's encoder-decoder architecture as specified in 
\cite{vaswani_attention_2017} corresponds to the leftmost masking pattern as
specified in \ref{masking_patterns}. T5 uses this architecture and masking pattern
\cite{raffel_exploring_2020}. As previously mentioned, BERT uses an encoder
only architecture \cite{devlin_bert_2019} and uses a fully-visible masking 
pattern with a special classification token \cite{raffel_exploring_2020}. GPT on
the other hand, uses a decoder only architecture \cite{radford_improving_nodate}
and uses causal masking \cite{liu_generating_2018}.

@@

@@

TODO: remake, text not visible
\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.1]{img/sparse_transformer.png}
  \caption{
    \cite{child_generating_2019}
  }
\end{figure}

@@

TODO: talk about masking the attention layers somewhere, was talked about in
the t5 paper
TODO: talk about masking, t5 paper lists this as the biggest differentiating factor
between models. the t5 paper is also good source for general comparison between different
architectures and approaches
(the above two are the same)

TODO: talk about the exact differences between fine-tuning,
i think bert modifies the weights and gpt only adds an extra layer after

TODO: elmo for feature based?

TODO: talk about GPT sparse transformer, used in gpt-3

@@

In the unsupervised pre-training phase, the model is given a set of long form text
\cite{radford_improving_nodate}. The supervised fine-tuning is used to train the
model for a specific task.

@@

BERT also uses unsupervised pre-training and supervised fine-tuning as the general training setup \cite{devlin_bert_2019}.
The largest amount of work and time is spent pre-training the model on a large set of unstructured
text. This teaches the model general language understanding. Fine-tuning can then be used to 
train the model for specific natural language tasks leveraging the existing language knowledge 
from the pre-training stage. Fine-tuning is usually a more lightweight operation than pre-training.

@@

\section{Problems}
TODO: write about hallucination, overconfidence, radicalization etc.
For example, when asking ChatGPT, which uses a slightly modified version of GPT-3 called 
GPT-3.5 [citation needed], to do a simple calculation, it answers confidently. 
With a prompt of "\texttt{what is 491*397}" we receive a response of 
"\texttt{The product of 491 and 397 is 195127.}". Unfortunately, using a calculator
leads to a different answer: 194927. Dangerously close, but not quite there.

>>>>

% TODO: temporary for having the "for the reader" section start from 0
\setcounter{section}{-1}

@@

\begin{abstract}
  This paper discusses three language models based on the transformer architecture.
  I aim decipher some differences between the models which seem very similar at first glance.
  The paper also takes a look at some applications and extensions to basic language
  models and some problems present in large language models, including biases
  and hallucination.
\end{abstract}

\section{For the reader}
There's a bit of length left in this paper for some extra explanations,
examples and images, so if there's anything specific you would like to
read about, please indicate that in your review. It can be quite hard to 
gauge the level of difficulty of different topics after being submerged 
in them for a while. Please keep in mind that I should not overlap too much
with specific language model applications nor the general transformer
architecture, since those are their own distinct topics assigned to 
other people.

@@

\subsection{Attention masking}
Attention masking prevents the model from attending to some tokens in the input. At any
given time $j$, the model is allowed to see input up to index $i$. Masking is used to prevent the model from
attending to future input which would trivialize training. \cite{raffel_exploring_2020}

The original transformer's encoder-decoder architecture as specified in 
\cite{vaswani_attention_2017} corresponds to the leftmost masking pattern as
specified in figure \ref{masking_patterns}. T5 uses this architecture and masking pattern
\cite{raffel_exploring_2020}. As previously mentioned, BERT uses an encoder
only architecture \cite{devlin_bert_2019} and uses a fully-visible masking 
pattern with a special classification token \cite{raffel_exploring_2020}. GPT on
the other hand, uses a decoder only architecture \cite{radford_improving_nodate}
and uses causal masking \cite{liu_generating_2018}.

@@

Calculating the entire attention history for a long sequence can be an expensive operation.
It is possible to use a sparse version of the attention masking pattern and still achieve 
comparable performance to a full attention mask. A sparse transformer with a fixed attention
pattern includes cells that can summarize the previous locations and propagate 
that forward. This allows computing longer sequences of input tokens and is used in GPT-3
\cite{brown_language_2020}. An illustrating example of different masking patterns
can be seen in figure \ref{sparse_transformer}. \cite{child_generating_2019}

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.1]{img/sparse_transformer.png}
  \caption{
    Sparse Transformer. Visualizing the attention masking pattern on a 6x6 image
    for two attention heads. On the left, the original transformer. In the middle,
    a sparse transformer using striding and on the right, a fixed sparse transformer.
    \cite{child_generating_2019}
  }
\label{sparse_transformer}
\end{figure}

@@

@@

In the unsupervised pre-training phase, the model is given a set of long form text. 
The supervised fine-tuning is used to train the
model for a specific task. In the supervised fine-tuning phase, an additional linear
output layer is added
and the pre-trained model's activation layer and linear output layer are trained
to the specific task. \cite{radford_improving_nodate}

@@

BERT also uses unsupervised pre-training and supervised fine-tuning as the general training setup \cite{devlin_bert_2019}.
The largest amount of work and time is spent pre-training the model on a large set of unstructured
text. This teaches the model general language understanding. Fine-tuning can then be used to 
train the model for specific natural language tasks leveraging the existing language knowledge 
from the pre-training stage. The pre-trained model is used to initialize a new model and 
then all the parameters are trained for a specific task. Fine-tuning is usually a more 
lightweight operation than pre-training. \cite{devlin_bert_2019}

@@

\section{Problems}
\subsection{Biases}
The writers of \cite{brown_language_2020} did a review of gender, racial and religious
biases in the model they presented in the paper. Their conclusion was that there were
certain biases in place which matched the biases present in the training data, that being
the internet.

Most highly educated and physical labor occupations had a male bias and in general, males
had a bias for being mentioned more often both in positive and negative contexts. Racial
sentiment depended on the size of the model, but in general, the sentiment for white,
black and middle eastern people was around neutral and for latinx, asian and indian it was
positive. For religion, violence was linked to islam at a greater rate than for other
major world religions.

\subsection{Hallucination}
Large languge models produce convincing text that can rival human written text \cite{brown_language_2020}.
It is possible to get these models to produce content which is 
not entirely factual or even completely nonsensical \cite{ji_survey_2023}.

For example, when asking ChatGPT, which uses a slightly modified version of GPT-3 called 
GPT-3.5 [citation needed], to do a simple calculation, it answers confidently. 
With a prompt of "\texttt{what is 491*397}" we receive a response of 
"\texttt{The product of 491 and 397 is 195127.}". Unfortunately, using a calculator
leads to a different answer: 194927. Dangerously close, but not quite there.
